---
title: "Anomaly Detection with R"
author: "Kelvin Kilel"
date: "2/7/2022"
output: html_document
---

# Research Question

I have been requested to check whether there are any anomalies in the given sales dataset. The objective of this task being fraud detection.

# Reading our dataset

```{r}
# Loading the reduired Libraries

library(tidyverse)
library(anomalize)

```

```{r}
# Loading our dataset
link <- 'http://bit.ly/CarreFourSalesDataset'
data <- read_csv(link)

```

```{r}
# Previewing the top records

head(data)

```

# Tidying the Dataset


```{r}
# Checking for missing values
colSums(is.na(data))

```

```{r}
# Converting the date column
data$Date <- as.Date(data$Date, format = "%m/%d/%Y")

```

```{r}
# Changing the dataset to tibble
#data$Date <- as.Date(data$Date, format = "%m%d%Y")

new_data <- as.tibble(data)

```

```{r}
is_tibble(new_data)

```

```{r}
# Aggregating sales to get daily records

new_data.anomaly <- aggregate(new_data["Sales"], by=new_data["Date"],sum)
head(new_data.anomaly)

```

# Anomaly Detection

```{r}
# Converting to tibble
new_data.anomaly <- as.tibble(new_data.anomaly)
is_tibble(new_data.anomaly)
```

```{r}
anomaly.detect <- new_data.anomaly %>%
  time_decompose(Sales, method = "stl", frequency = "auto", trend = "auto") %>%
  anomalize(remainder, method = "gesd", alpha = 0.05, max_anoms = 0.2) %>%
  plot_anomaly_decomposition()

```

```{r}
anomaly.detect

```

# Conclusion

There are anomalies as from the month of February. Carrefour Supermarket management should investigate why there are anomalies as from February.